Parenting Observation Experiment 2 Lexical Diversity Analysis

### 2019-01-10 parenting_obs_e1/parenting_obs_lexdiv.Rmd
lexical diversity analyses and figures

---
title: "Parenting observation lexical diversity"
author: "Vivian, George, and Emily"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

@@

```{r}
data <- read.csv("lexdiv_data.csv", header =TRUE)
#dem <- read.csv("parenting_proj_emilyhembacher_demo2016.csv", header =TRUE) # anonymized demographics data
#dem_anon = dem[,c("SID","age","gender","ethnicity","english","parent_ed","num_lang","Exclude")]
#write.csv(dem_anon, file="anonymized_demographics_parenting_obs_e1.csv")
dem <- read.csv("anonymized_demographics_parenting_obs_e1.csv", header=T)
conditions <- read.csv("joint_attention/conditions.csv")
load("paq/paq_demo.RData")

@@

```{r}
d <- left_join(data, conditions)%>%
  left_join(dem)%>%
  transmute(sid=SID, 
            types = Type.count, 
            tokens = Token.count, 
            lexdiv = Lexical.diversity, 
            condition= Condition, 
            video = Video, 
            age = age,
            gender = gender,
            parent_ed = parent_ed)%>%
  left_join(ids)%>%
  filter(!is.na(condition))
```

@@

##Lexical Diversity
```{r}
ms_lex <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "lexdiv") # Type/Token Ratio

lex_means <- d%>%
  group_by(condition)%>%
  summarise(mean = mean(lexdiv))

lex_sds <- d%>%
  group_by(condition)%>%
  summarise(sd = sd(lexdiv))

@@

ggplot(ms_lex, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Type/Token Ratio (TTR)") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="e1_TTR.pdf", width=3, height=3)
```

@@

ggplot(ms_tok, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Tokens") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="e1_word_tokens.pdf", width=3, height=3)
```

@@

ggplot(ms_tok, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Types") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few()
ggsave(file="e1_word_types.pdf", width=3, height=3)
```

>>>>

---
title: "Parenting observation lexical diversity - Exp 1"
author: "Vivian, George, and Emily"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

@@

```{r}
#data <- read.csv("lexdiv_data.csv", header=TRUE) # Emily's TTR calculation
data <- read.csv("MTLD_HDD.csv", header=TRUE) # Vivian's new TTR, MTLD, and HDD pipeline (generated by get_all_parent_speech.py)
data$SID = data$sid
data$sid = NULL
data$X = NULL
#dem <- read.csv("parenting_proj_emilyhembacher_demo2016.csv", header =TRUE) # anonymized demographics data
#dem_anon = dem[,c("SID","age","gender","ethnicity","english","parent_ed","num_lang","Exclude")]
#write.csv(dem_anon, file="anonymized_demographics_parenting_obs_e1.csv")
dem <- read.csv("anonymized_demographics_parenting_obs_e1.csv", header=T)
conditions <- read.csv("joint_attention/conditions.csv")
load("paq/paq_demo.RData") # do we use this??

@@

```{r}
d <- left_join(data, conditions)%>%
  left_join(dem)%>%
  transmute(sid=SID, 
            types = types,
            tokens = tokens, 
            TTR = TTR, 
            MTLD = MTLD,
            condition = Condition, 
            video = Video, 
            age = age,
            gender = gender,
            parent_ed = parent_ed)%>%
  left_join(ids)%>%
  filter(!is.na(condition))
```

@@

##Lexical Diversity
### TTR
```{r}
ms_lex <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "TTR") # Type/Token Ratio

lex_means <- d%>%
  group_by(condition)%>%
  summarise(mean = mean(TTR), sd = sd(TTR))

# report table with lex_means lex_sds

@@

ggplot(ms_lex, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Type/Token Ratio (TTR)") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="figs/e1_TTR.pdf", width=3, height=3)
```

###MTLD
```{r}
ms_lex_mtld <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "MTLD")

lex_means <- d%>%
  group_by(condition)%>%
  summarise(mean = mean(MTLD), sd = sd(MTLD))

# report table with lex_means lex_sds

ms_lex_mtld$condition <- factor(ms_lex_mtld$condition,
levels = c("con", "exp"),
labels = c("Control", "Video"))

ggplot(ms_lex_mtld, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("MTLD") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="figs/e1_MTLD.pdf", width=3, height=3)
```

@@

ggplot(ms_tok, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Tokens") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="figs/e1_word_tokens.pdf", width=3, height=3)
```

@@

ggplot(ms_tok, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Types") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few()
ggsave(file="figs/e1_word_types.pdf", width=3, height=3)
```

### 2019-01-10 parenting_obs_e2/Lexical_Diversity_Analysis.Rmd
lexical diversity analyses and figures

---
title: "Compare Lexical Diversity Measurements: MTLD and HD-D
author: "Vivian"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

@@

```{r}
data <- read.csv("MTLD_HDD.csv", header =TRUE) # generated by gen_
# dem <- read.csv("parenting_proj_emilyhembacher_demo2016.csv", header =TRUE)
conditions <- read.csv("conditions.csv")
```

@@

##TTR
```{r}
ms_lex <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "TTR")

lex_means <- d%>%
  group_by(condition)%>%
  summarise(mean = mean(TTR))

lex_sds <- d%>%
  group_by(condition)%>%
  summarise(sd = sd(TTR))

ms_lex$condition <- factor(ms_lex$condition,
levels = c("con", "exp"),
labels = c("control", "video"))

ggplot(ms_lex, aes(x = condition, y = mean, fill = condition)) + 
  ylim(0.0, 0.6)+
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("TTR") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="e2_TTR.pdf", width=3, height=3)
```
## MTLD
```{r}
ms_lex <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "MTLD")

lex_means <- d%>%
  group_by(condition)%>%
  summarise(mean = mean(MTLD))

lex_sds <- d%>%
  group_by(condition)%>%
  summarise(sd = sd(MTLD))

ms_lex$condition <- factor(ms_lex$condition,
levels = c("con", "exp"),
labels = c("control", "video"))

ggplot(ms_lex, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("MTLD") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() + 
  theme(legend.title = element_text(size=18), 
        legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), 
        axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.y = element_text(size=18))
```
## HDD
```{r}
ms_lex <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "HDD")

@@

ms_lex$condition <- factor(ms_lex$condition,
levels = c("con", "exp"),
labels = c("control", "video"))

ggplot(ms_lex, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("HD-D") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() + 
  theme(legend.title = element_text(size=18), 
        legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), 
        axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.y = element_text(size=18))
```
##Word Tokens
```{r}
ms_tok <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "tokens")

ms_tok$condition <- factor(ms_tok$condition,
levels = c("con", "exp"),
labels = c("control", "video"))

ggplot(ms_tok, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Tokens") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() + 
  theme(legend.title = element_text(size=18), 
        legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), 
        axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.y = element_text(size=18))
```

##Word Types
```{r}
ms_type <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "types")

ms_type$condition <- factor(ms_type$condition,
levels = c("con", "exp"),
labels = c("control", "video"))

ggplot(ms_type, aes(x = condition, y = mean, fill = condition)) +
  ylim(0, 100)+
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Types") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() + 
  theme(legend.title = element_text(size=18), 
        legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), 
        axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.y = element_text(size=18))
```

##Sentences Length
```{r}
ms_length <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "len_sentence")

ms_length$condition <- factor(ms_length$condition,
levels = c("con", "exp"),
labels = c("control", "video"))

ggplot(ms_length, aes(x = condition, y = mean, fill = condition)) +
  ylim(0,4.5)+
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Average Sentence Length") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() + 
  theme(legend.title = element_text(size=18), 
        legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), 
        axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.y = element_text(size=18))
```
##Sentences Length by videos
```{r}
ms_length <- d %>%
  group_by(video) %>%
  multi_boot_standard(col = "len_sentence")

@@

ggplot(ms_length, aes(x = video, y = mean, fill = video)) +
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Videos") + 
  ylab("Average Sentence Length") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() + 
  theme(legend.title = element_text(size=18), 
        legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), 
        axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.y = element_text(size=18))
```
##Sentences Length by age
```{r}
ms_length <- d %>%
  group_by(age) %>%
  multi_boot_standard(col = "len_sentence")

ms_length$age <- factor(ms_length$age,
levels = c("1", "2"),
labels = c("12-18month", "18-24month"))

ggplot(ms_length, aes(x = age, y = mean, fill = age)) +
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Age") + 
  ylab("Average Sentence Length") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() + 
  theme(legend.title = element_text(size=18), 
        legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), 
        axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.y = element_text(size=18))
```

@@

cor(d$ttr, d$mtld)
cor(d$ttr, d$hdd)
cor(d$mtld, d$hdd)

@@

#Conclusions
In McCarthy-Jarvis2010 paper, MTLD and HD-D has a correlation of r = .800; 
MTLD and TTR r =.322;
HD-D and TTR r = -.051 (N = 1584). 
In our result, MTLD and HD-D are the best correlated, although r = .621; MTLD and ttr has r = .543, which is higher than what M-J2010 reported.

All of the measure show that control group has higher lexical diversity than experimental group, with MTLD showing the biggest contrast and HD-D smallest.

>>>>

---
title: "Parenting Observation Experiment 2 Lexical Diversity Analysis"
author: "Vivian & George"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

@@

```{r}
data <- read.csv("MTLD_HDD.csv", header =TRUE) # generated by get_all_parent_speech.py
# dem <- read.csv("parenting_proj_emilyhembacher_demo2016.csv", header =TRUE)
conditions <- read.csv("conditions.csv")
```

@@

##TTR
```{r echo=FALSE}
ms_lex <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "TTR")

lex_mean_TTR <- d%>%
  group_by(condition)%>%
  summarise(mean = mean(TTR), sd = sd(TTR))

ms_lex$condition <- factor(ms_lex$condition,
levels = c("con", "exp"),
labels = c("Control", "Video"))

ggplot(ms_lex, aes(x = condition, y = mean, fill = condition)) + 
  ylim(0.0, 0.6)+
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("TTR") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="figs/e2_TTR.pdf", width=3, height=3)
```
## MTLD
```{r echo=FALSE}
ms_lex <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "MTLD")

lex_mean_MTLD <- d%>%
  group_by(condition)%>%
  summarise(mean = mean(MTLD), sd=sd(MTLD))

ms_lex$condition <- factor(ms_lex$condition,
levels = c("con", "exp"),
labels = c("Control", "Video"))

ggplot(ms_lex, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("MTLD") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="figs/e2_MTLD.pdf", width=3, height=3)
```
## HD-D
```{r echo=FALSE}
ms_lex <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "HDD")

@@

ms_lex$condition <- factor(ms_lex$condition,
levels = c("con", "exp"),
labels = c("Control", "Video"))

ggplot(ms_lex, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("HD-D") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few()
ggsave(file="figs/e2_HDD.pdf", width=3, height=3)
```
##Word Tokens
```{r echo=FALSE}
ms_tok <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "tokens")

ms_tok$condition <- factor(ms_tok$condition,
levels = c("con", "exp"),
labels = c("Control", "Video"))

ggplot(ms_tok, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Tokens") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="figs/e2_tokens.pdf", width=3, height=3)
```

##Word Types
```{r echo=FALSE}
ms_type <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "types")

ms_type$condition <- factor(ms_type$condition,
levels = c("con", "exp"),
labels = c("Control", "Video"))

ggplot(ms_type, aes(x = condition, y = mean, fill = condition)) +
  ylim(0, 100)+
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Types") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
ggsave(file="figs/e2_types.pdf", width=3, height=3)
```

##Sentence Length
```{r echo=FALSE}
ms_length <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "len_sentence")

ms_length$condition <- factor(ms_length$condition,
levels = c("con", "exp"),
labels = c("Control", "Video"))

ggplot(ms_length, aes(x = condition, y = mean, fill = condition)) +
  ylim(0,4.5)+
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Average Sentence Length") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few()
ggsave(file="figs/e2_sentence_length.pdf", width=3, height=3)
```

###Sentence Length by videos

```{r echo=FALSE}
ms_length <- d %>%
  group_by(video) %>%
  multi_boot_standard(col = "len_sentence")

@@

ggplot(ms_length, aes(x = video, y = mean, fill = video)) +
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Videos") + 
  ylab("Average Sentence Length") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() + 
  theme(legend.title = element_text(size=18), 
        legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), 
        axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.y = element_text(size=18))
```

###Sentence Length by age
(Maybe do a scatterplot with continuous scatterplot)

```{r echo=FALSE}
ms_length <- d %>%
  group_by(age) %>%
  multi_boot_standard(col = "len_sentence")

ms_length$age <- factor(ms_length$age,
levels = c("1", "2"),
labels = c("12-18", "18-24"))

ggplot(ms_length, aes(x = age, y = mean, fill = age)) +
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Age (months)") + 
  ylab("Average Sentence Length") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() + 
  theme(legend.title = element_text(size=18), 
        legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), 
        axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.y = element_text(size=18))
```

@@

round(cor(d[,c("TTR","MTLD","HDD","len_sentence")]), 3)

@@

#Conclusions
In McCarthy-Jarvis2010 paper, MTLD and HD-D has a correlation of r = .800; 
MTLD and TTR r =.322;
HD-D and TTR r = -.051 (N = 1584). 
In our results, MTLD and HD-D have the highest correlation (r = .714). Meanwhile, MTLD and TTR has r = .270, a bit lower than what M-J2010 reported. TTR and HDD are also modestly correlated (r = .426)

All of the measures suggest that the control group has higher lexical diversity than experimental group, with HD-D showing the biggest contrast and MTLD the smallest.
